The unphysicality of Hilbert spaces

### 2024-01-23 Articles/2023-HilbertQuantum/HilbertQuantum.tex
HilbertQuantum: removing change of orientation

More concretely, a change of variable $x$ to $y$ induces a unitary operator as it is just a change of representation. The transformation rules for wavefunctions simply follow by the fact that $\psi^\dagger(x) \psi(x) = \rho(x)$ gives us a probability density over $x$. Therefore we must have
\begin{equation}
	\begin{aligned}
		\rho(x) dx &= \rho(y) dy \\
		\rho(y) &= \rho(x) \frac{dx}{dy} \\
		\psi^\dagger(y) \psi(y) &= \psi^\dagger(x) \psi(x) \frac{dx}{dy}.
	\end{aligned}
\end{equation}
Therefore we can set 
\begin{equation}
	\begin{aligned}
		\psi(y) &= \psi(x(y)) \sqrt{\frac{dx}{dy}}.
	\end{aligned}
\end{equation}
If the change of variable does not change orientation, the above transformation is unitary.\footnote{If the orientation changes, the transformation is anti-unitary.}

\begin{prop}
	Let $\mathcal{H} = L^2(\mathbb{R})$ be the space of square integrable functions. Let $f : \mathbb{R} \to \mathbb{R}$ be an orientation-preserving diffeomorphism. Then the map $U : \mathcal{H} \to \mathcal{H}$ such that 
	$\phi(y) = U(\psi(x)) \mapsto \psi(f^{-1}(y)) \sqrt{\left(\frac{df}{dx}\right)^{-1}}$ is a unitary transformation.
\end{prop}
\begin{proof}
	Note that $f$ models a change of variables, therefore, in physics notation, $y = y(x) = f(x)$ and $x = x(y) = f^{-1}(y)$. Also note that both variables are real, therefore the Jacobian $\frac{df}{dx}$, which exists and  is nonzero because $f$ is a diffeomorphism, is real valued. Therefore $\left(\frac{df}{dx}\right)^\dagger = \frac{df}{dx}$. We have:
	\begin{equation}
		\begin{aligned}
			\int \phi^\dagger(y) \phi(y) dy &= \int U(\psi(x))^\dagger U(\psi(x)) dy \\
			 = \int \psi^\dagger(f^{-1}(y)) &\sqrt{\left(\frac{df}{dx}\right)^{-1}}\psi(f^{-1}(y)) \sqrt{\left(\frac{df}{dx}\right)^{-1}} dy \\
			 &= \int \psi^\dagger(f^{-1}(y)) \psi(f^{-1}(y)) \left(\frac{df}{dx}\right)^{-1} dy \\
			 &= \int \psi^\dagger(x(y)) \psi(x(y)) \frac{dx}{dy} dy \\
			 &= \int \psi^\dagger(x) \psi(x) dx.
		\end{aligned}
	\end{equation}
	The map $U$ preserved the inner product and therefore is a unitary transformation.
\end{proof}

>>>>

More concretely, a change of variable $x$ to $y$ induces a unitary operator as it is just a change of representation. The transformation rules for wavefunctions simply follow by the fact that $\psi^\dagger(x) \psi(x) = \rho(x)$ gives us a probability density over $x$. Therefore we must have
\begin{equation}
	\begin{aligned}
		\rho(x) dx &= \rho(y) dy \\
		\rho(y) &= \rho(x) \frac{dx}{dy} \\
		\psi^\dagger(y) \psi(y) &= \psi^\dagger(x) \psi(x) \frac{dx}{dy}.
	\end{aligned}
\end{equation}
Therefore we can set 
\begin{equation}
	\begin{aligned}
		\psi(y) &= \psi(x(y)) \sqrt{\frac{dx}{dy}}.
	\end{aligned}
\end{equation}
The above transformation is unitary.

\begin{prop}
	Let $\mathcal{H} = L^2(\mathbb{R})$ be the space of square integrable functions. Let $f : \mathbb{R} \to \mathbb{R}$ be an orientation-preserving diffeomorphism. Then the map $U : \mathcal{H} \to \mathcal{H}$ such that 
	$\phi(y) = U(\psi(x)) \mapsto \psi(f^{-1}(y)) \sqrt{\left|\frac{df}{dx}\right|^{-1}}$ is a unitary transformation.
\end{prop}
\begin{proof}
	Note that $f$ models a change of variables, therefore, in physics notation, $y = y(x) = f(x)$ and $x = x(y) = f^{-1}(y)$. Also note that both variables are real, therefore the Jacobian $\frac{df}{dx}$, which exists and  is nonzero because $f$ is a diffeomorphism, is real valued. Therefore $\left(\frac{df}{dx}\right)^\dagger = \frac{df}{dx}$. We have:
	\begin{equation}
		\begin{aligned}
			\int \phi^\dagger(y) \phi(y) dy &= \int U(\psi(x))^\dagger U(\psi(x)) dy \\
			 = \int \psi^\dagger(f^{-1}(y)) &\sqrt{\left|\frac{df}{dx}\right|^{-1}}\psi(f^{-1}(y)) \sqrt{\left|\frac{df}{dx}\right|^{-1}} dy \\
			 &= \int \psi^\dagger(f^{-1}(y)) \psi(f^{-1}(y)) \left|\frac{df}{dx}\right|^{-1} dy \\
			 &= \int \psi^\dagger(x(y)) \psi(x(y)) \frac{dx}{dy} dy \\
			 &= \int \psi^\dagger(x) \psi(x) dx.
		\end{aligned}
	\end{equation}
	The map $U$ preserved the inner product and therefore is a unitary transformation.
\end{proof}
